---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STA 506 2.0 Linear Regression Analysis

### Tutorial 1


A medical research team wants to study the relationship between the **head circumferences of newborns (centimetres) - HC** and the **gestational age (weeks) - GA**. A simple random sample of 57 babies was selected from the records of babies born in a certain hospital, and their head circumferences (centimetres) and gestational age (weeks) were recorded.  

In their attempt to find the relationship between head circumferences and gestational age, they plotted the head circumference versus the gestational age and obtained the following graph 

```{r hc, comment=NA, echo=FALSE, fig.cap="The scatter plot  of head circumference versus the gestational age. The Pearson correlation coeffifient is 0.99", warning=FALSE, message=FALSE, fig.height=3, fig.width=3}
set.seed(2020)
library(tidyverse)
GA <- c(rep(22, 4), rep(23.5, 5), rep(24, 4), 24, rep(25.5, 5),  rep(26, 4), rep(27, 3), rep(28, 4), rep(29, 5),
        rep(30.5, 3), rep(31, 5), rep(32.2, 5), rep(33.5, 5), rep(34, 4))
HC <- 3.91 + (0.78 * GA) + rnorm(50, sd=0.1)
df1 <- data.frame(GA=GA, HC=HC)
ggplot(df1, aes(x=GA, y=HC)) + geom_point() +
  coord_fixed(ratio=1.2) + xlab("Gestational age (weeks) - GA") + 
  ylab("Head circumference (cm) - HC") 

```

i) Comment on the scatter plot given above.

ii) Write the model you would fit to these data. Define all terms in it and state any assumptions you make regarding the model.

A simple linear regression analysis was performed with these data and the following outputs were obtained using R software.

**Output A**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
model <- lm(HC~GA)
model
```



\newpage

**Output B**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
summary(model)
```

**Output C**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
anova(model)
```


**Output D**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE, fig.height=3, fig.width=6}
library(broom)
fitmodel <- augment(model)
fitmodel$residmodify <- fitmodel$.resid
 fitmodel$residmodify <- ifelse(fitmodel$residmodify >0.11, 0.1, fitmodel$residmodify)
  fitmodel$residmodify <- ifelse(fitmodel$residmodify < -0.22, -0.1, fitmodel$residmodify)
p1 <- qplot(data=fitmodel, y=residmodify, x=.fitted)+ ggtitle("Figure a") + ylab("Residuals")
p2 <- ggplot(fitmodel, 
       aes(sample=.resid))+
  stat_qq() + 
  stat_qq_line() +
  labs(x="Theoretical Quantiles",
       y="Sample Quantiles") + ggtitle("Figure b")
library(patchwork)
p1|p2
```

\newpage


**Output E**

```{r, comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
shapiro.test(fitmodel$.resid)
```

iii) Write the fitted regression model.


iv) Complete the ANOVA table below (Copy the table below in your answer script and complete it.).

\begin{table}[!h]
\center
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Source of variation & DF & Sum of squares (SS) & Mean Square (MS) & F-value &  p-value\\ \hline
Regression &  &  &  &  &  \\ \hline
Residual error &  &  &  &  &  \\ \hline
Total &  &  &  &  &  \\ \hline
\end{tabular}
\end{table}

v) Test the significance of the model fitted. You should clearly write the hypothesis, decision and conclusions.

vi) What proportion of the variation in the response is explained by the model fitted?

vii) Two undergraduates studying Biostatistics were looking at this analysis.

\begin{enumerate}
\item[ ]  Student A: said that the results strongly suggest that this model is highly significant and can be used for prediction purposes. 
\item[ ]  Student B: said that the results show the fitted model is not appropriate for this case and this model cannot be used for prediction. 
\end{enumerate}


\begin{enumerate}
\item[] With whom would you agree? Justify your argument using the results given above (Outputs A-E).  
\end{enumerate}


viii) The point on the graph (Figure 2, below), coloured in red, represents a baby with gestational age 31 weeks and head circumference 28.03 cm, (GA=31, HC=28.03). 

```{r hc2, comment=NA, echo=FALSE, fig.cap="The scatterplot of head circumference versus the gestational age. The point, (GA=31, HC=28.03), is coloured in red on the graph.", warning=FALSE, message=FALSE, fig.height=4, fig.width=4}
df2 = df1[42, ]
ggplot(df1, aes(x=GA, y=HC)) + geom_point() +
  coord_fixed(ratio=1.2) + xlab("Gestational age (weeks) - GA") + 
  ylab("Head circumference (cm) - HC") + geom_point(data=df2, aes(x=GA, y=HC), colour="red", shape=4, lwd=5)
```

\begin{enumerate}
\item[a. ]  Find the fitted value for this point. 
\item[b. ]  Find the residual for this point.
\end{enumerate}

